---
title: "Vessel Traffic Data Analysis"
author: "Shannon Ricci"
date: "3/21/2019"
output: html_document
---

## Project Description



## Vessel Traffic Data Preparation
Vessel traffic data was downloaded from <https://marinecadastre.gov/ais/> on 3/1/2019 for Zone 17 in year 2017. The area of interest for this analysis is the area surrounding the lower Florida Keys, USA which includes several MPAs in the Florida Keys National Marine Sanctuary. The original AIS data will be filtered to include just points within the bounds of the AOI. These csv files will then be used to convert the point data into track lines using the Marine Cadastre Track Builder Tool for ArcPro.

```{r filter.AIS }
library(data.table)
#ran out of memory when tried to do this in a loop. Manually filter and save to csv for AIS data by month.

ais.filelist<-list.files("/Users/sbrown/Documents/research/vesseltraffic", pattern=glob2rx("*AIS*.csv$"), full.names = T)
  
ais.zone17.jan2017<-fread(ais.filelist[1])
ais.zone17.jan2017<-ais.zone17.jan2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.jan2017,"ais_fknms_jan2017.csv")

ais.zone17.feb2017<-fread(ais.filelist[2])
ais.zone17.feb2017<-ais.zone17.feb2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.feb2017,"ais_fknms_feb2017.csv")

ais.zone17.mar2017<-fread(ais.filelist[3])
ais.zone17.mar2017<-ais.zone17.mar2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.mar2017,"ais_fknms_mar2017.csv")

ais.zone17.apr2017<-fread(ais.filelist[4])
ais.zone17.apr2017<-ais.zone17.apr2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.apr2017,"ais_fknms_apr2017.csv")

ais.zone17.may2017<-fread(ais.filelist[5])
ais.zone17.may2017<-ais.zone17.may2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.may2017,"ais_fknms_may2017.csv")

ais.zone17.jun2017<-fread(ais.filelist[6])
ais.zone17.jun2017<-ais.zone17.jun2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.jun2017,"ais_fknms_jun2017.csv")

ais.zone17.jul2017<-fread(ais.filelist[7])
ais.zone17.jul2017<-ais.zone17.jul2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.jul2017,"ais_fknms_jul2017.csv")

ais.zone17.aug2017<-fread(ais.filelist[8])
ais.zone17.aug2017<-ais.zone17.aug2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.aug2017,"ais_fknms_aug2017.csv")

ais.zone17.sep2017<-fread(ais.filelist[9])
ais.zone17.sep2017<-ais.zone17.sep2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.sep2017,"ais_fknms_sep2017.csv")

ais.zone17.oct2017<-fread(ais.filelist[10])
ais.zone17.oct2017<-ais.zone17.oct2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.oct2017,"ais_fknms_oct2017.csv")

ais.zone17.nov2017<-fread(ais.filelist[11])
ais.zone17.nov2017<-ais.zone17.nov2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.nov2017,"ais_fknms_nov2017.csv")

ais.zone17.dec2017<-fread(ais.filelist[12])
ais.zone17.dec2017<-ais.zone17.dec2017[LAT %between% c(24.25,24.75) & LON %between% c(-82.75, -80.75),]
write.csv(ais.zone17.dec2017,"ais_fknms_dec2017.csv")

```

## Add Vessel Type based on Vessel Codes
Was unable to use the join tool in Arc Pro to merge vessel codes with the vessel type description. After creating tracks from the point csv files in Arc Pro (using the Track Builder tool), the shape files for the track lines for each month were exported. In this next step, the shapefiles are loaded into R, and then vessel types are added to the shapefiles based on vessel code.

```{r process.tracks.func, warning=FALSE}

process.tracks<-function(track.file,vessel.codes,output.name){
  tracks_tmp<-st_read(track.file)
  tracks_coded<-merge(tracks_tmp,vessel.codes,by.x="VesselType",by.y="vessel_code")
  tracks_coded$MMSI<-as.character(tracks_coded$MMSI)
  setwd("C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017_coded")
  st_write(tracks_coded,output.name)
  
  
  tracks_dt<-as.data.table(tracks_coded)
  tracks_dt[Length==0,"Length"]<-NA #set vessel lengths of 0 to NA
  
  tracks_summary<-tracks_dt[,.(count=.N, avg.length=mean(Length,na.rm=T)),by="vessel_group"]
  tracks_unique<-tracks_dt[,.(count=.N),by=c("MMSI","vessel_group","VesselName")][order(-count),]
  
  result<-list(tracks_summary=tracks_summary,tracks_unique=tracks_unique)
  return(result)

}

```

```{r code.tracks, warning=FALSE}
library(sf)
library(data.table)
library(ggplot2)

vessel.codes<-fread("C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/VesselTypeCodes2018.csv")

#setwd("C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017_coded")

#jan2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_jan2017.shp"
output.name="tracks_jan2017c.shp"
tracks_jan2017<-process.tracks(track.file,vessel.codes,output.name)

#feb2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_feb2017.shp"
output.name="tracks_feb2017c.shp"
tracks_feb2017<-process.tracks(track.file,vessel.codes,output.name)

#mar2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_mar2017.shp"
output.name="tracks_mar2017c.shp"
tracks_mar2017<-process.tracks(track.file,vessel.codes,output.name)

#apr2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_apr2017.shp"
output.name="tracks_apr2017c.shp"
tracks_apr2017<-process.tracks(track.file,vessel.codes,output.name)

#may2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_may2017.shp"
output.name="tracks_may2017c.shp"
tracks_may2017<-process.tracks(track.file,vessel.codes,output.name)

#jun2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_jun2017.shp"
output.name="tracks_jun2017c.shp"
tracks_jun2017<-process.tracks(track.file,vessel.codes,output.name)

#jul2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_jul2017.shp"
output.name="tracks_jul2017c.shp"
tracks_jul2017<-process.tracks(track.file,vessel.codes,output.name)

#aug2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_aug2017.shp"
output.name="tracks_aug2017c.shp"
tracks_aug2017<-process.tracks(track.file,vessel.codes,output.name)

#sep2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_sep2017.shp"
output.name="tracks_sep2017c.shp"
tracks_sep2017<-process.tracks(track.file,vessel.codes,output.name)

#oct2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_oct2017.shp"
output.name="tracks_oct2017c.shp"
tracks_oct2017<-process.tracks(track.file,vessel.codes,output.name)

#nov2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_nov2017.shp"
output.name="tracks_nov2017c.shp"
tracks_nov2017<-process.tracks(track.file,vessel.codes,output.name)

#dec2017
track.file="C:/Users/swbrown/Documents/vessel_noise_project/vessel_analysis/tracks_fknms_2017/tracks_fknms_dec2017.shp"
output.name="tracks_dec2017c.shp"
tracks_dec2017<-process.tracks(track.file,vessel.codes,output.name)

```


Track lines from each vessel category that best represented areas of high activity for that vessel category were extracted. In ArcPro, points were generated along each of these track lines spaced 1 km apart. Those points were exported from ArcPro as csv files so that we can generate the time between each of those points based on average vessel speed. The following function accomplishes this task. The final output is a csv file with point position information (in UTM), the depth of that source point, time between the two points, and the number of sections. This is the format and data required to import a moving source line into dBSea.
```{r generate.tracks.function}
generate.tracks<-function(point.file,speed,draft,output.name){
  track.points<-fread(point.file)
  track.points<-st_as_sf(track.points,coords = c('POINT_X','POINT_Y'), crs= "+init=epsg:32617")
  
  point.dist<-rep(NA,length(track.points$OBJECTID))
  for (i in seq(from=1, to=length(track.points$OBJECTID)-1,by= 1)){
    point.dist[i+1]<-st_distance(track.points[i+1,],track.points[i,])
  }
  point.dist[1]<-0
  
  point.time<-rep(NA,length(track.points$OBJECTID))
  for (i in seq(from=1, to=length(track.points$OBJECTID)-1,by= 1)){
    point.time[i+1]<-point.dist[i+1]/speed
  }
  point.time[1]<-0
  
  depth<-rep(draft,length(point.dist))
  sections<-rep(0,length(point.dist))
  
  tmp.track<-data.table(st_coordinates(track.points),depth,point.time,sections)
  colnames(tmp.track)<-c("Easting","Northing","Depth","Time","Sections")
  
  write.csv(tmp.track,output.name,row.names = F, quote = F)
}

```

```{r generate.tracks}
library(data.table)
library(sf)

#set vessel specific parameters:
# speed <- #in m/s
# draft <- #in m
# output.name<- #path with .csv file name

#pcsail satori
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_satori.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_satori_track.csv"
generate.tracks(file,5.35,1,output.name) #speed is vessel cat average

#pcsail pelican
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_pelican.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_pelican_track.csv"
generate.tracks(file,3.02,1,output.name)

#pcsail blue pearl
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_bluepearl.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/pcsail_bluepearl_track.csv"
generate.tracks(file,5.35,1,output.name) #speed is vessel cat average

#fishing stimstar near
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_stimstar_near.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_stimstar_near_track.csv"
generate.tracks(file,12.86,1.5,output.name)

#fishing stimstar far
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_stimstar_far.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_stimstar_far_track.csv"
generate.tracks(file,12.86,1.5,output.name)

#fishing big daddy
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_bigdaddy.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/fishing_bigdaddy_track.csv"
generate.tracks(file,3.85,1.5,output.name)

#passenger capt blinky
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_captblinky.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_captblinky_track.csv"
generate.tracks(file,5.6,1,output.name)

#passenger empress of the seas
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_empressoftheseas.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_empressoftheseas_track.csv"
generate.tracks(file,6.36,7.1,output.name)

#passenger grand nellie (sailing)
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_grandnellie.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_grandnellie_track.csv"
generate.tracks(file,2.67,1,output.name)

#passenger yankee freedom iii
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_yankeefreedom.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/passenger_yankeefreedom_track.csv"
generate.tracks(file,12.65,2,output.name)

#tank overseas new york
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tank_overseasnewyork.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tank_overseasnewyork_track.csv"
generate.tracks(file,7.77,12.2,output.name)

#tug genesis patriot
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tug_genesispatriot.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tug_genesispatriot_track.csv"
generate.tracks(file,4.32,5.1,output.name)

#tug kirsten grace
file<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tug_kirstengrace.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_fknms/tracks_dbsea/tug_kirstengrace_track.csv"
generate.tracks(file,4.6,3,output.name) #speed is vessel cat average


```

Styles for Figures
Colors:
Tank: teal (#1b9e77)
Cargo: orange (#d95f02)
Fishing: green (#66a61e)
Passenger: purple (#7570b3)
Tug: yellow (#e6ab02)
Pleasure Craft/Sailing: pink (#e7298a)
Other: gray (#666666)


Using R code provided by Veirs et al 2016, extract median source level spectra for the vessel catergories of interest for this study. Data downloaded from http://www.beamreach.org/data/staff-research/ship-noise/
The following code produced: a csv file that was used to load source level information into dBSea noise propagation modeling software and a figure of the source level spectra for each vessel type (using code adapted from Veirs et al)
```{r, eval = F}
library(data.table)
library(pracma)
library(scales)

load("data/vessel_SL/df_ALL_quant_linear_SL.Rdata")

dt_vessel_SL<-data.table(df_ALL_quant_linear_SL)

#interpolate SL at octave frequencies for dBSea input
library(pracma)
fcenter  <- 10^3 * (2 ^ (seq(-14,17,1)/3))
fcenter_dBsea <- 10^3 * (2 ^ (seq(-18,22,1)/3))

tug<-dt_vessel_SL[class == "Tug", .(q50.avg = mean(q50)),by=fhz]
tug_octave<-interp1(tug$fhz,tug$q50.avg,fcenter)
#tug_octave<-rbind(fcenter,tug_octave)

cargo<-dt_vessel_SL[class == "Cargo", .(q50.avg = mean(q50)),by=fhz]
cargo_octave<-interp1(cargo$fhz,cargo$q50.avg,fcenter)
#cargo_octave<-rbind(fcenter, cargo_octave)

tanker<-dt_vessel_SL[class == "Tanker", .(q50.avg = mean(q50)),by=fhz]
tanker_octave<-interp1(tanker$fhz,tanker$q50.avg,fcenter)
#tanker_octave<-rbind(fcenter, tanker_octave)

passenger<-dt_vessel_SL[class == "Passenger", .(q50.avg = mean(q50)),by=fhz]
passenger_octave<-interp1(passenger$fhz,passenger$q50.avg,fcenter)
#passenger_octave<-rbind(fcenter, passenger_octave)

fishing<-dt_vessel_SL[class == "Fishing", .(q50.avg = mean(q50)),by=fhz]
fishing_octave<-interp1(fishing$fhz,fishing$q50.avg,fcenter)
#fishing_octave<-rbind(fcenter, fishing_octave)

pleasure_craft<-dt_vessel_SL[class == "Pleasure craft", .(q50.avg = mean(q50)),by=fhz]
pleasure_craft_octave<-interp1(pleasure_craft$fhz,pleasure_craft$q50.avg,fcenter)
#pleasure_craft_octave<-rbind(fcenter, pleasure_craft_octave)

vessel_type_SL_octave<-data.frame(pleasure_craft_octave,fishing_octave,passenger_octave,tanker_octave,cargo_octave,tug_octave)

#vessel_type_SL<-t(vessel_type_SL_octave)
#write.csv(vessel_type_SL_octave,file = "vessel_type_SL.csv",row.names = T, quote = F)


vessel_type_SL<-dt_vessel_SL[class == "Tug" | class == "Pleasure craft" | class == "Fishing" | class == "Passenger" | class == "Cargo" | class == "Fishing" | class == "Tanker", .(q50.avg = mean(q50)),by=c("class","fhz")]

vessel_order<-c("Tanker","Cargo","Passenger","Tug","Fishing","Pleasure craft")

vessel_type_SL$class2<-factor(vessel_type_SL$class, levels = vessel_order)

SL_vessel_type_fig<-ggplot(vessel_type_SL)+
  scale_x_log10(breaks=c(10,100,1000,10000,100000),labels = trans_format("log10", math_format(10^.x)),lim=c(20,100000)) +
  annotation_logticks(sides = "tb") +
  geom_line(aes(x=fhz,y=q50.avg, group = class2, color = class2), size = 1) +##, linetype=class)) +
  # geom_smooth(aes(x=fhz,y=q50, group=class, color=class),se=FALSE,method = "loess", formula = y ~ x, degree=2, span=.005, size=1)+
  scale_y_continuous(breaks=seq(30,280,10),lim=c(100,160))+ 
  ylab(expression(paste("Source level (dB re 1 ", mu, "Pa"^2,"/Hz @ 1m)")))+
  xlab("Frequency (Hz)")+
  theme_bw()+
  theme(panel.grid.major = element_line(colour="black", size=0.10),panel.grid.minor.y=element_line(color="black",size=0.05))+
  scale_color_manual(name = "Vessel class", values = c("Tanker" = "#1b9e77","Cargo" = "#d95f02","Fishing" = "#66a61e","Passenger" = "#7570b3", "Tug" = "#e6ab02","Pleasure craft" = "#e7298a"))

ggsave("SL_vesseltype_fig.eps",SL_vessel_type_fig)
ggsave("/Users/swricci/Documents/gis714_geocomp/final_presentation/SL_vesseltype_fig.tiff",SL_vessel_type_fig, width = 6, height = 4, units = "in", dpi = 600)
```

```{r}
#generate sound speed profile figure
fknms_jul_speed<-fread("C:/Users/sbrown/Documents/research/fknms_july_speedprofile.csv")

speed_profile<-ggplot(fknms_jul_speed)+
  geom_line(aes(x = speed, y = depth), size = 1)+
  scale_y_reverse()+
  scale_x_continuous(breaks=seq(1480,1540,20),lim = c(1480,1550),position = "top")+
  labs(
    y = "Depth (m)",
    x = "Sound speed (m/s)")+
  theme_classic()

ggsave("C:/Users/sbrown/Documents/research/fknms_july_speedprofile.tiff",speed_profile, width = 2, height = 3, dpi = 600)

```

```{r}
library(ggplot2)
library(data.table)
library(dplyr)

tracks2017<-fread(paste(getwd(),"/data/tracks2017.csv",sep=""))
tracks2017[Length == 0, "Length"] <-NA 
tracks2017<-tracks2017[,`:=`(month=format(as.Date(TrckStr, '%m/%d/%Y'),'%m'))]

#vessel_types<-c("Not Available","Other","Cargo","Tanker","Tug Tow","Passenger","Fishing","Pleasure Craft/Sailing")
# 
# tracks2017.summary<-tracks2017[,.(count = .N, avg.length = mean(Length, na.rm=T), unique.vessels = length(unique(MMSI)), avg.speed = mean(SOG), median.speed  = median(SOG), avg.track.length = mean(Shp_Lng)), by="vssl_gr"]
# 
# tracks2017.summary.month<-tracks2017[,.(count = .N, avg.length = mean(Length, na.rm=T), unique.vessels = length(unique(MMSI)), avg.track.length = mean(Shp_Lng)), by=c("vssl_gr","month")]

tracks2017.summary.filt<-tracks2017[Shape_Length >=0.0045,.(count = .N, avg.length = mean(Length, na.rm=T), unique.vessels = length(unique(MMSI)), avg.speed = mean(SOG), median.speed  = median(SOG), avg.track.length = mean(Shp_Lng)), by="month"]


tracks2017.filt<-tracks2017[Shape_Length >=0.0045,]
tracks2017.filt<-tracks2017.filt%>%
  mutate(class = case_when(vssl_gr == "Cargo"  ~ "Cargo",
                           vssl_gr == "Tanker"  ~ "Tanker",
                           vssl_gr == "Passenger" ~ "Passenger",
                           vssl_gr == "Tug Tow" ~ "Tug/Tow",
                        vssl_gr == "Fishing" ~ "Fishing",
                        vssl_gr == "Pleasure Craft/Sailing" ~ "Recreational",
                        vssl_gr == "Other" | vssl_gr == "Not Available" ~ "Other"))

vessel_order<-c("Tanker","Cargo","Passenger","Tug/Tow","Fishing","Recreational","Other")

tracks2017.filt$class2<-factor(tracks2017.filt$class, levels = vessel_order)

tracks.bar<-ggplot(tracks2017.filt)+
  geom_bar(aes(x = month, fill = class2), position = "fill")+
  scale_fill_manual(name = "Vessel Class", values = c("Tanker" = "#1b9e77","Cargo" = "#d95f02","Fishing" = "#66a61e","Passenger" = "#7570b3", "Tug/Tow" = "#e6ab02","Recreational" = "#e7298a", "Other" = "#666666"))+
  scale_x_discrete(labels=c("01" = "Jan", "02" = "Feb","03" = "Mar", "04" = "Apr", "05" = "May","06" = "Jun", "07" = "Jul", "08" = "Aug", "09" = "Sep", "10" = "Oct","11" = "Nov", "12" = "Dec"))+
  scale_y_continuous(name = "Proportion of tracks")+
  theme_classic(base_size = 14)

ggsave("/Users/swricci/Documents/gis714_geocomp/final_presentation/tracks_barchart.tiff",tracks.bar, width = 6, height = 4, dpi = 600)

```