---
title: "FKNMS Mooring Buoy Vessel Noise"
author: "Shannon Ricci"
date: "5/7/2019"
output: html_document
---

```{r}





```


##Get sound spectra for dBSea
Using R code provided by Veirs et al 2016, extract median source level spectra for the vessel catergories of interest for this study. Data downloaded from <http://www.beamreach.org/data/staff-research/ship-noise/>

The following code produced: a csv file that was used to load source level information into dBSea noise propagation modeling software and a figure of the source level spectra for each vessel type (using code adapted from Veirs et al)
```{r, eval = F}
library(data.table)
library(pracma)
library(scales)

load("data/vessel_SL/df_ALL_quant_linear_SL.Rdata")

dt_vessel_SL<-data.table(df_ALL_quant_linear_SL)

#interpolate SL at octave frequencies for dBSea input
library(pracma)
fcenter  <- 10^3 * (2 ^ (seq(-14,17,1)/3))
fcenter_dBsea <- 10^3 * (2 ^ (seq(-18,22,1)/3))

recreational_q50<-dt_vessel_SL[class == "Pleasure craft", .(q50.avg = mean(q50)),by=fhz]
recreational_q50_octave<-interp1(recreational_q50$fhz,recreational_q50$q50.avg,fcenter)
recreational_q50_dbsea<-rbind(fcenter, recreational_q50_octave)

recreational_q95<-dt_vessel_SL[class == "Pleasure craft", .(q95.avg = mean(q95)),by=fhz]
recreational_q95_octave<-interp1(recreational_q95$fhz,recreational_q95$q95.avg,fcenter)
recreational_q95_dbsea<-rbind(fcenter, recreational_q95_octave)

write.csv(recreational_q50_dbsea,file = "data/recreational_q50_SL.csv",row.names = T, quote = F)

write.csv(recreational_q95_dbsea,file = "data/recreational_q95_SL.csv",row.names = T, quote = F)

recreational_SL<-zone.tracks %>% 
  gather(key = "class", value = "tracks",`Tanker`:`Other`)

recreational_SL<-dt_vessel_SL[class == "Pleasure craft", .(q50.avg = mean(q50), q95.avg = mean(q95)),by=fhz]

recreational_SL<-recreational_SL %>% 
  gather(key = "quantile", value = "SPL",`q50.avg`:`q95.avg`)


ggplot() +
  scale_x_log10(breaks=c(10,100,1000,10000,100000),labels = trans_format("log10", math_format(10^.x)),lim=c(20,100000)) +
  annotation_logticks(sides = "tb") +
  geom_line(data=recreational_SL,aes(x = fhz, y = SPL, group = quantile, color = quantile), size = 1)+
scale_y_continuous(breaks=seq(30,280,10),lim=c(100,160))+ 
  ylab(expression(paste("Source level (dB re 1 ", mu, "Pa"^2,"/Hz @ 1m)")))+
  xlab("Frequency (Hz)")+
  theme_bw()+
  theme(panel.grid.major = element_line(colour="black", size=0.10),panel.grid.minor.y=element_line(color="black",size=0.05))+
  theme(legend.position = "right")+
  scale_color_manual(values = c("q50.avg" = "black","q95.avg" = "red"))

```

```{r}
library(data.table)
library(tidyr)
library(ggplot2)


speed_profile <- fread("C:/Users/sbrown/Documents/research/speedprofiles.csv")

speed_profile<-speed_profile  %>%
  gather(key = "time", value = "soundspeed",`annual`:`july`)

ggplot(speed_profile)+
  geom_line(aes(x = soundspeed, y = depth, group = time, color = time),size = 1)+
  scale_y_reverse()+
  scale_x_continuous(breaks = seq(1480,1540,10),position = "top")+
  theme_classic()


```

```{r generate.tracks.function}
generate.tracks<-function(point.file,speed,draft,output.name){
  track.points<-fread(point.file)
  track.points<-st_as_sf(track.points,coords = c('POINT_X','POINT_Y'), crs= "+init=epsg:32617")
  
  point.dist<-rep(NA,length(track.points$OBJECTID))
  for (i in seq(from=1, to=length(track.points$OBJECTID)-1,by= 1)){
    point.dist[i+1]<-st_distance(track.points[i+1,],track.points[i,])
  }
  point.dist[1]<-0
  
  point.time<-rep(NA,length(track.points$OBJECTID))
  for (i in seq(from=1, to=length(track.points$OBJECTID)-1,by= 1)){
    point.time[i+1]<-point.dist[i+1]/speed
  }
  point.time[1]<-0
  
  depth<-rep(draft,length(point.dist))
  sections<-rep(0,length(point.dist))
  
  tmp.track<-data.table(st_coordinates(track.points),depth,point.time,sections)
  colnames(tmp.track)<-c("Easting","Northing","Depth","Time","Sections")
  
  write.csv(tmp.track,output.name,row.names = F, quote = F)
}
```

Use function to generate tracks for each representative vessel:
```{r generate.tracks}
library(data.table)
library(sf)

#set vessel specific parameters:
# speed <- #in m/s
# draft <- #in m
# output.name<- #path with .csv file name

#WDR
#neverland
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wdr_neverland_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wdr_neverland.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wdr_familytime_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wdr_familytime.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational cat average

#SK
#godspeed
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/sk_godspeed_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/sk_godspeed.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational cat average

#marianne
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/sk_marianne_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/sk_marianne.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational cat average

#orion
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/sk_orion_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/sk_orion.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational cat average

#NFS
#neverland
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/nfs_neverland_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/nfs_neverland.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#wheel estate
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/nfs_wheelestate_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/nfs_wheelestate.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#WSB
#flying fish
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wsb_flyingfish_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wsb_flyingfish.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#gratitude
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wsb_gratitude_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wsb_gratitude.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#neverland
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wsb_neverland_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wsb_neverland.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#osprey moon
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/wsb_ospreymoon_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/wsb_ospreymoon.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#LKSPA
#ospreymoon
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/lkspa_ospreymoon_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/lkspa_ospreymoon.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#flyingcolors
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/lkspa_flyingcolors_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/lkspa_flyingcolors.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#questeria
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/lkspa_questeria_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/lkspa_questeria.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#SOM
#kathrynann
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/som_kathrynann_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/som_kathrynann.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#bandulu
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/som_bandulu_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/som_bandulu.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

#elizabear
file<-"C:/Users/sbrown/Documents/dBSea_mbuoy/track_points/som_elizabear_points.csv"
output.name<-"C:/Users/sbrown/Documents/dBSea_mbuoy/tracks_dbsea/som_elizabear.csv"
generate.tracks(file,5.35,1,output.name) #speed is recreational average

```
